---
title: "README"
author: "Mathew Ling, Steph de Silva, Maddie Davey, Nicholas Tierney, Adam Gruer"
date: "27/10/2017"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ozflights

Welcome to Ozflights! Get the data for passenger, aircraft and international freight movements for both metropolitan and regional airports in Australia. The dataset covers 1985-2016. 

## Installation

You can install `ozflights` from github:

```
#install.packages("devtools")
devtools::install_github("ropenscilabs/ozflights")
```

## Package Author's Notes

Data was available at URL as at 17/10/26. Data is imported into R and cleaned by removing redundant headers and transforming into a tidy format.

Airport ranking variable was removed: the interested party can recalculate.

Uses the annual airport traffic data from the Australian Government Department of Infrastructure and Regional Development located at: https://bitre.gov.au/publications/ongoing/airport_traffic_data.aspx

Contact details for the data itself can be found in the original spreadsheet, but please don’t contact the aviation stats team for problems with the package!

## Airline Passengers

To access the data on airport passengers, use the function `airport_passengers()`.
```{r, fig.show='hold'}
passengers <- ozflights::airport_passengers()
```

We'll just look at the total for Australia. 

```{r test-figure, fig.show='hold'}
library(ggplot2)
passengers_aus <- dplyr::filter(passengers, airport == "TOTAL AUSTRALIA")
passengers_aus$domest <- factor(passengers_aus$domest, levels = c(TRUE, FALSE), labels = c("Domestic", "International"))
passengers_aus_in <- dplyr::filter(passengers_aus, type == "inbound")
ggplot(passengers_aus_in)+
  labs(x="Year", y="Number of Passengers")+ 
  facet_grid(domest~., scale="free")+
  geom_line(aes(year, count, colour = type))+
  theme_light()+
  ggtitle("Inbound Passengers")
```


